---
title: "Exploratory Data Analysis"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r,include = FALSE}
library(tidyverse)
library(plotly)
library(purrr)
```

```{r, include = FALSE}
hap_df = read_csv("./data/combined_happiness.csv")|>
  mutate(
    year = as.factor(year)
  )

regional_hap_df = hap_df|>
  group_by(region,year) |>
  summarise(
    average_gdp_per_capita = mean(gdp_per_capita, na.rm = TRUE),
    average_social_support = mean(social_support, na.rm = TRUE),
    average_healthy_life_expectancy = mean(health_life_expectancy, na.rm = TRUE),
    average_freedom_to_make_life_choices = mean(freedom_to_make_life_choices, na.rm = TRUE),
    average_perception_of_corruption = mean(perception_of_corruption, na.rm = TRUE),
    average_generosity = mean(generosity, na.rm = TRUE)
  )|>
  pivot_longer(
    average_gdp_per_capita:average_generosity,
    names_to ="factor",
    names_prefix = "average_",
    values_to = "average_score"
  )

```

# Explore Six Factors

## Which country has the highest and lowest score for the six factors
### GDP Per Captia
```{r}
hap_df |>
  mutate(country = fct_reorder(country, gdp_per_capita))|>
  plot_ly (
  x = ~country, y = ~gdp_per_capita, color = ~year,
  type = "bar", colors = "viridis") |>
  layout(
    title = "GDP Per Capita",
    xaxis = list(
      title = "country",
      tickfont = list(size = 6)
      ),
    yaxis = list(title = "GDP Per Capita"),
    autosize = F, width = 1000, height = 500
  )
```

### Social Support
```{r}
hap_df |>
  mutate(country = fct_reorder(country, social_support))|>
  plot_ly (
  x = ~country, y = ~social_support, color = ~year,
  type = "bar", colors = "viridis") |>
  layout(
    title = "Social Support/Having Someone to Count On In Time of Trouble",
    xaxis = list(
      title = "country",
      tickfont = list(size = 6)
      ),
    yaxis = list(title = "Score"),
    autosize = F, width = 1000, height = 500
  )
```

### life Expectancy
```{r}
hap_df |>
  mutate(country = fct_reorder(country, health_life_expectancy))|>
  plot_ly (
  x = ~country, y = ~health_life_expectancy, color = ~year,
  type = "bar", colors = "viridis") |>
  layout(
    title = "Healthy Life Expectancy",
    xaxis = list(
      title = "country",
      tickfont = list(size = 6)
      ),
    yaxis = list(title = "Score"),
    autosize = F, width = 1000, height = 500
  )
```

### Freedom to make Choice 
```{r}
hap_df |>
  mutate(
    country = fct_reorder(country, freedom_to_make_life_choices))|>
  plot_ly (
  x = ~country, y = ~freedom_to_make_life_choices, color = ~year,
  type = "bar", colors = "viridis") |>
  layout(
    title = "Freedom to Make Life Choices",
    xaxis = list(
      title = "country",
      tickfont = list(size = 6)
      ),
    yaxis = list(title = "Score"),
    autosize = F, width = 1000, height = 500
  )
```

### Perception of Corruption
```{r}
hap_df |>
  mutate(country = fct_reorder(country, perception_of_corruption))|>
  plot_ly (
  x = ~country, y = ~perception_of_corruption, color = ~year,
  type = "bar", colors = "viridis") |>
  layout(
    title = "Perception of Corruption",
    xaxis = list(
      title = "country",
      tickfont = list(size = 6)
      ),
    yaxis = list(title = "Score"),
    autosize = F, width = 1000, height = 500
  )
```

### Generosity
```{r}
hap_df |>
  mutate(country = fct_reorder(country, generosity))|>
  plot_ly (
  x = ~country, y = ~generosity, color = ~year,
  type = "bar", colors = "viridis") |>
  layout(
    title = "Genersoity",
    xaxis = list(
      title = "country",
      tickfont = list(size = 6)
      ),
    yaxis = list(title = "Score"),
    autosize = F, width = 1000, height = 500
  )
```

### Iteration
```{r, warning=FALSE}
factor_country_level <- function(data, variable, title) {
  p = data %>%
    mutate(country = fct_reorder(country, .data[[variable]])) %>%
    plot_ly(
      x = ~country, 
      y = as.formula(paste0("~", variable)), 
      color = ~year,
      type = 'bar', 
      colors = 'viridis'
    ) %>%
    layout(
      title = title,
      xaxis = list(title = "Country", tickfont = list(size = 6)),
      yaxis = list(title = "Score"),
      autosize = F, width = 1000, height = 500
    )
  
  print(p)
}

# Create a list of variables and corresponding titles
variables_and_titles <- list(
  list(variable = "gdp_per_capita", title = "GDP Per Capita"),
  list(variable = "social_support", title = "Social Support/Having Someone to Count On In Time of Trouble"),
  list(variable = "health_life_expectancy", title = "Healthy Life Expectancy"),
  list(variable = "freedom_to_make_life_choices", title = "Freedom to Make Life Choices"),
  list(variable = "perception_of_corruption", title = "Perception of Corruption"),
  list(variable = "generosity", title = "Generosity")
)

# Use map to create a list of plots
plots <- map(variables_and_titles, ~factor_country_level(hap_df, .x$variable, .x$title))

for (plot in plots) {
  print(plot)
}
```

## World Trends of the Six Happiness Indicators from 2015 to 2019
```{r}
hap_df |>
  group_by(year) |>
  summarise(
    average_gdp_per_capita = mean(gdp_per_capita, na.rm = TRUE),
    average_social_support = mean(social_support, na.rm = TRUE),
    average_healthy_life_expectancy = mean(health_life_expectancy, na.rm = TRUE),
    average_freedom_to_make_life_choices = mean(freedom_to_make_life_choices, na.rm = TRUE),
    average_perception_of_corruption = mean(perception_of_corruption, na.rm = TRUE),
    average_generosity = mean(generosity, na.rm = TRUE)
  )|>
  pivot_longer(
    average_gdp_per_capita:average_generosity,
    names_to ="factor",
    names_prefix = "average_",
    values_to = "average_score"
  )|>
  plot_ly(
    x = ~year, y = ~average_score, color = ~factor,
    type = "scatter", mode = "lines+markers", colors = "viridis") |>
  layout(
    title = "World Trends of Happiness Indicators from 2015 to 2019",
    xaxis = list( title = "Year"),
    yaxis = list(title = "Average Score"),
    autosize = F, width = 1000, height = 500
  )
```


## Which region has the highest and lowest score for the six factors
### GDP Per Capita
```{r}
hap_df |>
  mutate(region = fct_reorder(region, gdp_per_capita))|>
  group_by(region, year)|>
  summarise(
    average_gdp_per_capita = mean(gdp_per_capita, na.rm = TRUE)
  ) |>
  plot_ly (
  x = ~region, y = ~average_gdp_per_capita, color = ~year,
  type = "bar", colors = "viridis") |>
  layout(
    title = "GDP Per Capita By Region",
    xaxis = list(title = "Region"),
    yaxis = list(title = "Average Score"),
    autosize = F, width = 1000, height = 500
  )
```

**Regional Trend**
```{r}
regional_hap_df |>
  filter(factor == "gdp_per_capita")|>
  plot_ly(
      x = ~year, y = ~average_score, color = ~region,
      mode = 'lines+markers', type = 'scatter', colors = 'viridis',
      name = ~region) |>
    layout(
      title = paste("Regional Trend for GDP Per Captia from 2015 to 2019"),
      xaxis = list(title = "Year"),
      yaxis = list(title = "Average Score"),
      autosize = F, width = 1000, height = 500
    )
```

### Social Support
```{r}
hap_df |>
  mutate(region = fct_reorder(region, social_support))|>
  group_by(region, year)|>
  summarise(
    average_social_support = mean(social_support, na.rm = TRUE)
  ) |>
  plot_ly (
  x = ~region, y = ~average_social_support, color = ~year,
  type = "bar", colors = "viridis") |>
  layout(
    title = "Social Support By Region",
    xaxis = list(title = "Region"),
    yaxis = list(title = "Average Score"),
    autosize = F, width = 1000, height = 500
  )
```

**Regional Trend**
```{r}
regional_hap_df |>
  filter(factor == "social_support")|>
  plot_ly(
      x = ~year, y = ~average_score, color = ~region,
      mode = 'lines+markers', type = 'scatter', colors = 'viridis',
      name = ~region) |>
    layout(
      title = paste("Regional Trend for Social Support from 2015 to 2019"),
      xaxis = list(title = "Year"),
      yaxis = list(title = "Average Score"),
      autosize = F, width = 1000, height = 500
    )
```

### Life Expectancy
```{r}
hap_df |>
  mutate(region = fct_reorder(region, health_life_expectancy))|>
  group_by(region, year)|>
  summarise(
    average_life_expectancy = mean(health_life_expectancy, na.rm = TRUE)
  ) |>
  plot_ly (
  x = ~region, y = ~average_life_expectancy, color = ~year,
  type = "bar", colors = "viridis") |>
  layout(
    title = "Healthy Life Expectancy By Region",
    xaxis = list(title = "Region"),
    yaxis = list(title = "Average Score"),
    autosize = F, width = 1000, height = 500
  )
```

**Regional Trend**
```{r}
regional_hap_df |>
  filter(factor == "healthy_life_expectancy")|>
  plot_ly(
      x = ~year, y = ~average_score, color = ~region,
      mode = 'lines+markers', type = 'scatter', colors = 'viridis',
      name = ~region) |>
    layout(
      title = paste("Regional Trend for Life Expectancy from 2015 to 2019"),
      xaxis = list(title = "Year"),
      yaxis = list(title = "Average Score"),
      autosize = F, width = 1000, height = 500
    )
```


### Freedom to Make Choice
```{r}
hap_df |>
  mutate(region = fct_reorder(region, freedom_to_make_life_choices))|>
  group_by(region, year)|>
  summarise(
    average_freedom = mean(freedom_to_make_life_choices, na.rm = TRUE)
  ) |>
  plot_ly (
  x = ~region, y = ~average_freedom, color = ~year,
  type = "bar", colors = "viridis") |>
  layout(
    title = "Freedom to Make Choice By Region",
    xaxis = list(title = "Region"),
    yaxis = list(title = "Average Score"),
    autosize = F, width = 1000, height = 500
  )
```

**Regional Trend**
```{r}
regional_hap_df |>
  filter(factor == "freedom_to_make_life_choices")|>
  plot_ly(
      x = ~year, y = ~average_score, color = ~region,
      mode = 'lines+markers', type = 'scatter', colors = 'viridis',
      name = ~region) |>
    layout(
      title = paste("Regional Trend for Freedom to Make Life Choices from 2015 to 2019"),
      xaxis = list(title = "Year"),
      yaxis = list(title = "Average Score"),
      autosize = F, width = 1000, height = 500
    )
```

### Perception of Corruption
```{r}
hap_df |>
  mutate(region = fct_reorder(region, perception_of_corruption))|>
  group_by(region, year)|>
  summarise(
    average_corruption = mean(perception_of_corruption, na.rm = TRUE)
  ) |>
  plot_ly (
  x = ~region, y = ~average_corruption, color = ~year,
  type = "bar", colors = "viridis") |>
  layout(
    title = "Perception of Corruption By Region",
    xaxis = list(title = "Region"),
    yaxis = list(title = "Average Score"),
    autosize = F, width = 1000, height = 500
  )
```

**Regional Trend**
```{r}
regional_hap_df |>
  filter(factor == "perception_of_corruption")|>
  plot_ly(
      x = ~year, y = ~average_score, color = ~region,
      mode = 'lines+markers', type = 'scatter', colors = 'viridis',
      name = ~region) |>
    layout(
      title = paste("Regional Trend for Perception of Corruption from 2015 to 2019"),
      xaxis = list(title = "Year"),
      yaxis = list(title = "Average Score"),
      autosize = F, width = 1000, height = 500
    )
```

### Generosity
```{r}
hap_df |>
  mutate(region = fct_reorder(region, generosity))|>
  group_by(region, year)|>
  summarise(
    average_generosity = mean(generosity, na.rm = TRUE)
  ) |>
  plot_ly (
  x = ~region, y = ~average_generosity, color = ~year,
  type = "bar", colors = "viridis") |>
  layout(
    title = "Generosity By Region",
    xaxis = list(title = "Region"),
    yaxis = list(title = "Average Score"),
    autosize = F, width = 1000, height = 500
  )
```

**Regional Trend**
```{r}
regional_hap_df |>
  filter(factor == "generosity")|>
  plot_ly(
      x = ~year, y = ~average_score, color = ~region,
      mode = 'lines+markers', type = 'scatter', colors = 'viridis',
      name = ~region) |>
    layout(
      title = paste("Regional Trend for Generosity from 2015 to 2019"),
      xaxis = list(title = "Year"),
      yaxis = list(title = "Average Score"),
      autosize = F, width = 1000, height = 500
    )
```

## World Map
```{r include=FALSE}
library(ggplot2)
library(stringr)
library(maps)
library(magrittr)
library(rvest)
library(countrycode)
```

```{r}
iso_codes = countrycode::codelist[, c("un.name.en", "iso3c")]
names(iso_codes) = c("Country", "ISO3")
```

```{r, warning = FALSE, message = FALSE}
world_data <- ggplot2::map_data('world')
world_data <- fortify(world_data)
```

```{r}
hap_df['ISO3'] <- iso_codes$ISO3[match(hap_df$country, iso_codes$Country)]
world_data["ISO3"] <- iso_codes$ISO3[match(world_data$region, iso_codes$Country)]
```

```{r}
hap_df = hap_df |>
  mutate(
    ISO3 = if_else(country == "United States", "USA", ISO3),
    ISO3 = if_else(country == "United Kingdom", "GBR", ISO3),
    ISO3 = if_else(country == "Venezuela", "VEN", ISO3),
    ISO3 = if_else(country == "Czech Republic", "CZE", ISO3),
    ISO3 = if_else(country == "Taiwan", "TWN", ISO3),
    ISO3 = if_else(country == "South Korea", "KOR", ISO3),
    ISO3 = if_else(country == "Bolivia", "BOL", ISO3),
    ISO3 = if_else(country == "Moldova", "MDA", ISO3),
    ISO3 = if_else(country == "Russia", "RUS", ISO3),
    ISO3 = if_else(country == "Hong Kong", "CHN", ISO3),
    ISO3 = if_else(country == "Vietnam", "VNM", ISO3),
    ISO3 = if_else(country == "Turkey", "TUR", ISO3),
    ISO3 = if_else(country == "Macedonia", "MKD", ISO3),
    ISO3 = if_else(country == "Laos", "LAO", ISO3),
    ISO3 = if_else(country == "Swaziland", "SWZ", ISO3),
    ISO3 = if_else(country == "Palestinian Territories", "PSE", ISO3),
    ISO3 = if_else(country == "Iran", "IRN", ISO3),
    ISO3 = if_else(country == "Congo (Kinshasa)", "COD", ISO3),
    ISO3 = if_else(country == "Congo (Brazzaville)", "COG", ISO3),
    ISO3 = if_else(country == "Tanzania", "TZA", ISO3),
    ISO3 = if_else(country == "Syria", "SYR", ISO3),
    ISO3 = if_else(country == "Puerto Rico", "PRI", ISO3)
  )|>
  mutate(country = recode(str_trim(country),
                         "United States" = "USA",
                         "United Kingdom" = "UK",
                         "Congo (Kinshasa)" = "Democratic Republic of the Congo",
                         "Congo (Brazzaville)" = "Republic of Congo"))
```

```{r}
world_data = world_data|>
  mutate(country = region)|>
  select(-region)

worldjoin <- inner_join(world_data, hap_df, by = "country")
```

```{r, fig.width=12, fig.height=9, fig.retina=2}
## compile all map theme configurations
cleanup <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank(),
  panel.background = element_rect(fill = "white"),
  plot.title = element_text(hjust = 0.5)
)

## plot our merged data

gdp <- worldjoin %>% 
  ggplot(mapping = aes(
    x=long,
    y=lat,
    group=group)) +
  scale_fill_viridis_c(option = "C") +
  coord_fixed(1.3) +
  geom_polygon(aes(fill=gdp_per_capita)) +
  ggtitle("GDP Per Capita") +
  cleanup

gdp
```


# Explore Happiness Score
```{r include=FALSE}
combined_happiness = read.csv("./combined_happiness.csv") |>
  janitor::clean_names()

combined_happiness$region[combined_happiness$region == "Eastern Asian"] <- "Eastern Asia"
```


## Part 1: Which country has the highest and lowest happiness score?

In order to explore the distribution of happiness score all over the world, we draw bar charts of happiness score for each year in descending order showing different region with color and legends. We also use heat map in world map to show the geographic distribution of happiness score over the world.

### 2015
#### Bar Chat
```{r}
combined_happiness |>
  filter(year == "2015") |>
  arrange(desc(score)) |>
  ggplot(aes(x = reorder(country, -score), y = score, fill = region)) +
  geom_bar(stat = "identity") +
  labs(title = "2015 World Happiness Score Rank", x = "Countries", y = "Happiness Score") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 4)) +
  theme(legend.position = "bottom")
```

In year 2015, Switzerland has the highest happiness score while Togo has the lowest happiness score among 158 countries.

#### World Map
```{r}
hp_2015 <- worldjoin |>
  filter(year == "2015") |>
  ggplot(mapping = aes(
    x=long,
    y=lat,
    group=group)) +
  scale_fill_viridis_c(option = "C") +
  coord_fixed(1.3) +
  geom_polygon(aes(fill=score)) +
  ggtitle("Happiness Score 2015") +
  cleanup

hp_2015
```

In general, countries in Europe, Australia, North and South America have higher happiness score while countries in Africa and Middle East Asia has relatively lower happiness score in 2015.

### 2016
#### Bar Chart
```{r}
combined_happiness |>
  filter(year == "2016") |>
  arrange(desc(score)) |>
  ggplot(aes(x = reorder(country, -score), y = score, fill = region)) +
  geom_bar(stat = "identity") +
  labs(title = "2016 World Happiness Score Rank", x = "Countries", y = "Happiness Score") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 4)) +
  theme(legend.position = "bottom")
```

In year 2016, Denmark has the highest happiness score while Burundi has the lowest happiness score among 157 countries.

#### World Map
```{r}
hp_2016 <- worldjoin |>
  filter(year == "2016") |>
  ggplot(mapping = aes(
    x=long,
    y=lat,
    group=group)) +
  scale_fill_viridis_c(option = "C") +
  coord_fixed(1.3) +
  geom_polygon(aes(fill=score)) +
  ggtitle("Happiness Score 2016") +
  cleanup

hp_2016
```

In general, countries in Europe, Australia, North and South America have higher happiness score while countries in Africa and Middle East Asia has relatively lower happiness score in 2016.

### 2017
#### Bar Chart
```{r}
combined_happiness |>
  filter(year == "2017") |>
  arrange(desc(score)) |>
  ggplot(aes(x = reorder(country, -score), y = score, fill = region)) +
  geom_bar(stat = "identity") +
  labs(title = "2017 World Happiness Score Rank", x = "Countries", y = "Happiness Score") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 4)) +
  theme(legend.position = "bottom")
```

In year 2017, Norway has the highest happiness score while Central African Republic has the lowest happiness score among 155 countries.

#### World Map
```{r}
hp_2017 <- worldjoin |>
  filter(year == "2017") |>
  ggplot(mapping = aes(
    x=long,
    y=lat,
    group=group)) +
  scale_fill_viridis_c(option = "C") +
  coord_fixed(1.3) +
  geom_polygon(aes(fill=score)) +
  ggtitle("Happiness Score 2017") +
  cleanup

hp_2017
```

In general, countries in Europe, Australia, North and South America have higher happiness score while countries in Africa and Middle East Asia has relatively lower happiness score in 2017.

### 2018
#### Bar Chart
```{r}
combined_happiness |>
  filter(year == "2018") |>
  arrange(desc(score)) |>
  ggplot(aes(x = reorder(country, -score), y = score, fill = region)) +
  geom_bar(stat = "identity") +
  labs(title = "2018 World Happiness Score Rank", x = "Countries", y = "Happiness Score") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 4)) +
  theme(legend.position = "bottom")
```

In year 2018, Finland has the highest happiness score while Burundi has the lowest happiness score among 156 countries.

#### World Map
```{r}
hp_2018 <- worldjoin |>
  filter(year == "2018") |>
  ggplot(mapping = aes(
    x=long,
    y=lat,
    group=group)) +
  scale_fill_viridis_c(option = "C") +
  coord_fixed(1.3) +
  geom_polygon(aes(fill=score)) +
  ggtitle("Happiness Score 2018") +
  cleanup

hp_2018
```

In general, countries in Europe, Australia, North and South America have higher happiness score while countries in Africa and Middle East Asia has relatively lower happiness score in 2018.

### 2019
#### Bar Chart
```{r}
combined_happiness |>
  filter(year == "2019") |>
  arrange(desc(score)) |>
  ggplot(aes(x = reorder(country, -score), y = score, fill = region)) +
  geom_bar(stat = "identity") +
  labs(title = "2019 World Happiness Score Rank", x = "Countries", y = "Happiness Score") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 4)) +
  theme(legend.position = "bottom")
```

In year 2019, Finland has the highest happiness score while South Sudan has the lowest happiness score among 156 countries.

#### World Map
```{r}
hp_2019 <- worldjoin |>
  filter(year == "2019") |>
  ggplot(mapping = aes(
    x=long,
    y=lat,
    group=group)) +
  scale_fill_viridis_c(option = "C") +
  coord_fixed(1.3) +
  geom_polygon(aes(fill=score)) +
  ggtitle("Happiness Score 2019") +
  cleanup

hp_2019
```

In general, countries in Europe, Australia, North and South America have higher happiness score while countries in Africa and Middle East Asia has relatively lower happiness score in 2019.

### Average in 5 years
#### Bar Chart
```{r}
hp_avg = 
combined_happiness |>
  select(year, country, region, score) |>
  pivot_wider(
    names_from = "year",
    values_from = "score"
  ) |>
  na.omit() |>
  pivot_longer(
    "2015":"2019",
    names_to = "year", 
    values_to = "score"
  ) |>
  group_by(year, country) |>
  mutate(score_avg = mean(score)) 

hp_avg|>
  arrange(desc(score_avg)) |>
  ggplot(aes(x = reorder(country, -score_avg), y = score_avg, fill = region)) +
  geom_bar(stat = "identity") +
  labs(title = "Average World Happiness Score Rank 2015-2019", x = "Countries", y = "Happiness Score") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 4)) +
  theme(legend.position = "bottom")
```

We only keep countries that are in the top rank for all five years. Among those 144 countries, Danmark has the highest average happiness score while Burundi has the lowest average happiness score in 5 years.

#### World Map
```{r}
wold_avg <- inner_join(world_data, hp_avg, by = "country")
hp_avg_plot <- wold_avg |>
  ggplot(mapping = aes(
    x=long,
    y=lat,
    group=group)) +
  scale_fill_viridis_c(option = "C") +
  coord_fixed(1.3) +
  geom_polygon(aes(fill=score_avg)) +
  ggtitle("Average Happiness Score in 5 years") +
  cleanup

hp_avg_plot
```

In general, countries in Europe, Australia, North and South America have higher average happiness score while countries in Africa and Middle East Asia has relatively lower average happiness score from 2015 to 2019.

The geographic distribution of World Happiness Score is stable from 2015 to 2019. European Countries such as Finland, Switzerland, and Danmark tend to have highest happiness score while African countires such as South Sudan and Central African Republic tend to have lowest happiness score. Also, countries in Europe, Australia, North and South America have higher happiness score while countries in Africa and Middle East Asia has relatively lower happiness score.


## Part 2: Which region has the highest and lowest happiness score?
### 2015
#### Bar Chart
```{r}
combined_happiness |>
  filter(year == "2015") |>
  group_by(region) |>
  mutate(region_avg = mean(score)) |>
  arrange(desc(region_avg)) |>
  summarize(region = unique(region), score = unique(region_avg)) |>
  ggplot(aes(x = reorder(region, -score), y = score, fill = region)) +
  geom_bar(stat = "identity") +
  labs(title = "2015 World Happiness Score Rank", x = "Regions", y = "Happiness Score") +
  scale_x_discrete(labels = label_wrap_gen(width = 12)) +
  theme(legend.position = "none")
```

In year 2015, Australia and New Zealand has the highest happiness score while Sub-Saharan Africa has the lowest happiness score among 10 regions.

### 2016
#### Bar Chart
```{r}
combined_happiness |>
  filter(year == "2016") |>
  group_by(region) |>
  mutate(region_avg = mean(score)) |>
  arrange(desc(region_avg)) |>
  summarize(region = unique(region), score = unique(region_avg)) |>
  ggplot(aes(x = reorder(region, -score), y = score, fill = region)) +
  geom_bar(stat = "identity") +
  labs(title = "2016 World Happiness Score Rank", x = "Regions", y = "Happiness Score") +
  scale_x_discrete(labels = label_wrap_gen(width = 12)) +
  theme(legend.position = "none")
```

In year 2016, Australia and New Zealand has the highest happiness score while Sub-Saharan Africa has the lowest happiness score among 10 regions.


### 2017
#### Bar Chart
```{r}
combined_happiness |>
  filter(year == "2017") |>
  group_by(region) |>
  mutate(region_avg = mean(score)) |>
  arrange(desc(region_avg)) |>
  summarize(region = unique(region), score = unique(region_avg)) |>
  ggplot(aes(x = reorder(region, -score), y = score, fill = region)) +
  geom_bar(stat = "identity") +
  labs(title = "2017 World Happiness Score Rank", x = "Regions", y = "Happiness Score") +
  scale_x_discrete(labels = label_wrap_gen(width = 12)) +
  theme(legend.position = "none")
```

In year 2017, Australia and New Zealand has the highest happiness score while Sub-Saharan Africa has the lowest happiness score among 10 regions.



### 2018
#### Bar Chart
```{r}
combined_happiness |>
  filter(year == "2018") |>
  group_by(region) |>
  mutate(region_avg = mean(score)) |>
  arrange(desc(region_avg)) |>
  summarize(region = unique(region), score = unique(region_avg)) |>
  ggplot(aes(x = reorder(region, -score), y = score, fill = region)) +
  geom_bar(stat = "identity") +
  labs(title = "2018 World Happiness Score Rank", x = "Regions", y = "Happiness Score") +
  scale_x_discrete(labels = label_wrap_gen(width = 12)) +
  theme(legend.position = "none")
```

In year 2018, Australia and New Zealand has the highest happiness score while Sub-Saharan Africa has the lowest happiness score among 10 regions.


### 2019
#### Bar Chart
```{r}
combined_happiness |>
  filter(year == "2019") |>
  group_by(region) |>
  mutate(region_avg = mean(score)) |>
  arrange(desc(region_avg)) |>
  summarize(region = unique(region), score = unique(region_avg)) |>
  ggplot(aes(x = reorder(region, -score), y = score, fill = region)) +
  geom_bar(stat = "identity") +
  labs(title = "2019 World Happiness Score Rank", x = "Regions", y = "Happiness Score") +
  scale_x_discrete(labels = label_wrap_gen(width = 12)) +
  theme(legend.position = "none")
```

In year 2019, Australia and New Zealand has the highest happiness score while Sub-Saharan Africa has the lowest happiness score among 10 regions.

### Average in 5 years
#### Bar Chart
```{r}
hp_avg|>
  group_by(region) |>
  mutate(region_avg = mean(score)) |>
  arrange(desc(score_avg)) |>
  summarize(region = unique(region), score = unique(region_avg)) |>
  ggplot(aes(x = reorder(region, -score), y = score, fill = region)) +
  geom_bar(stat = "identity") +
  labs(title = "Average World Happiness Score Rank 2015-2019", x = "Countries", y = "Happiness Score") +
  scale_x_discrete(labels = label_wrap_gen(width = 12)) +
  theme(legend.position = "none")
```

From 2015 to 2019, Australia and New Zealand has the highest average happiness score while Sub-Saharan Africa has the lowest average happiness score among 10 regions.


## Part 3: What are some trends of mean happiness score in different regions from 2015 – 2019?
### Scatter Plot
```{r}
combined_happiness |>
  group_by(year, region) |>
  mutate(region_avg = mean(score)) |>
  arrange(desc(region_avg)) |>
  summarize(region = unique(region), score = unique(region_avg)) |>
  ggplot(aes(x = year, y = score, color = region)) +
  geom_point() + 
  geom_line() +
  labs(title = "World Happiness Score Trend in 10 Regions", x = "year", y = "Happiness Score")


```

According to the scatter plot, from 2015 to 2019, Australia and New Zealand has stable happiness scores and it's the highest amnong 10 regions. North America has second highest happiness score among 10 regions with the highest happiness score in year 2015 and gradually decreasing trend. Western Europe has the third highest happiness score with an increasing trend in 5 years. Latin America and Caribbean has the fourth highest happiness score with a decreasing trend in 5 years. Eastern Asia has the fifth highest happiness score with a stable trend. Middle East and Northern Africa, Southeastern Asia, and Central and Eastern Europe have comparative happiness score from 2015 to 2017, but Central and Eastern Europe has a increasing trend and higher happiness score after year 2017 while the other two regions have a decreasing trend and lower happiness score after year 2017. Southern Asia and Sub-Saharan Africa has the second lowest and lowest happiness score in 5 years, respectively. However, Sub-Saharan Africa has shown an increasing trend while Southern Asia has shown an decreasing trend since year 2017.





### Social Support
```{r}
hap_df |>
  mutate(country = fct_reorder(country, social_support))|>
  plot_ly (
  x = ~country, y = ~social_support, color = ~year,
  type = "bar", colors = "viridis") |>
  layout(
    title = "Social Support/Having Someone to Count On In Time of Trouble",
    xaxis = list(
      title = "country",
      tickfont = list(size = 6)
      ),
    yaxis = list(title = "Score"),
    autosize = F, width = 1000, height = 500
  )
```

### Life Expectancy
```{r}
hap_df |>
  mutate(country = fct_reorder(country, health_life_expectancy))|>
  plot_ly (
  x = ~country, y = ~health_life_expectancy, color = ~year,
  type = "bar", colors = "viridis") |>
  layout(
    title = "Healthy Life Expectancy",
    xaxis = list(
      title = "country",
      tickfont = list(size = 6)
      ),
    yaxis = list(title = "Score"),
    autosize = F, width = 1000, height = 500
  )
```

### Freedom to Make Choice 
```{r}
hap_df |>
  mutate(
    country = fct_reorder(country, freedom_to_make_life_choices))|>
  plot_ly (
  x = ~country, y = ~freedom_to_make_life_choices, color = ~year,
  type = "bar", colors = "viridis") |>
  layout(
    title = "Freedom to Make Life Choices",
    xaxis = list(
      title = "country",
      tickfont = list(size = 6)
      ),
    yaxis = list(title = "Score"),
    autosize = F, width = 1000, height = 500
  )
```

### Perception of Corruption
```{r}
hap_df |>
  mutate(country = fct_reorder(country, perception_of_corruption))|>
  plot_ly (
  x = ~country, y = ~perception_of_corruption, color = ~year,
  type = "bar", colors = "viridis") |>
  layout(
    title = "Perception of Corruption",
    xaxis = list(
      title = "country",
      tickfont = list(size = 6)
      ),
    yaxis = list(title = "Score"),
    autosize = F, width = 1000, height = 500
  )
```

### Generosity
```{r}
hap_df |>
  mutate(country = fct_reorder(country, generosity))|>
  plot_ly (
  x = ~country, y = ~generosity, color = ~year,
  type = "bar", colors = "viridis") |>
  layout(
    title = "Genersoity",
    xaxis = list(
      title = "country",
      tickfont = list(size = 6)
      ),
    yaxis = list(title = "Score"),
    autosize = F, width = 1000, height = 500
  )
```



## World Trends of the Six Happiness Indicators from 2015 to 2019
```{r}
hap_df |>
  group_by(year) |>
  summarise(
    average_gdp_per_capita = mean(gdp_per_capita, na.rm = TRUE),
    average_social_support = mean(social_support, na.rm = TRUE),
    average_healthy_life_expectancy = mean(health_life_expectancy, na.rm = TRUE),
    average_freedom_to_make_life_choices = mean(freedom_to_make_life_choices, na.rm = TRUE),
    average_perception_of_corruption = mean(perception_of_corruption, na.rm = TRUE),
    average_generosity = mean(generosity, na.rm = TRUE)
  )|>
  pivot_longer(
    average_gdp_per_capita:average_generosity,
    names_to ="factor",
    names_prefix = "average_",
    values_to = "average_score"
  )|>
  plot_ly(
    x = ~year, y = ~average_score, color = ~factor,
    type = "scatter", mode = "lines+markers", colors = "viridis") |>
  layout(
    title = "World Trends of Happiness Indicators from 2015 to 2019",
    xaxis = list( title = "Year"),
    yaxis = list(title = "Average Score"),
    autosize = F, width = 1000, height = 500
  )
```


## Which region has the highest and lowest score for the six factors
### GDP Per Capita
```{r}
hap_df |>
  mutate(region = fct_reorder(region, gdp_per_capita))|>
  group_by(region, year)|>
  summarise(
    average_gdp_per_capita = mean(gdp_per_capita, na.rm = TRUE)
  ) |>
  plot_ly (
  x = ~region, y = ~average_gdp_per_capita, color = ~year,
  type = "bar", colors = "viridis") |>
  layout(
    title = "GDP Per Capita By Region",
    xaxis = list(title = "Region"),
    yaxis = list(title = "Average Score"),
    autosize = F, width = 1000, height = 500
  )
```

**Regional Trend**
```{r}
regional_hap_df |>
  filter(factor == "gdp_per_capita")|>
  plot_ly(
      x = ~year, y = ~average_score, color = ~region,
      mode = 'lines+markers', type = 'scatter', colors = 'viridis',
      name = ~region) |>
    layout(
      title = paste("Regional Trend for GDP Per Captia from 2015 to 2019"),
      xaxis = list(title = "Year"),
      yaxis = list(title = "Average Score"),
      autosize = F, width = 1000, height = 500
    )
```

### Social Support
```{r}
hap_df |>
  mutate(region = fct_reorder(region, social_support))|>
  group_by(region, year)|>
  summarise(
    average_social_support = mean(social_support, na.rm = TRUE)
  ) |>
  plot_ly (
  x = ~region, y = ~average_social_support, color = ~year,
  type = "bar", colors = "viridis") |>
  layout(
    title = "Social Support By Region",
    xaxis = list(title = "Region"),
    yaxis = list(title = "Average Score"),
    autosize = F, width = 1000, height = 500
  )
```

**Regional Trend**
```{r}
regional_hap_df |>
  filter(factor == "social_support")|>
  plot_ly(
      x = ~year, y = ~average_score, color = ~region,
      mode = 'lines+markers', type = 'scatter', colors = 'viridis',
      name = ~region) |>
    layout(
      title = paste("Regional Trend for Social Support from 2015 to 2019"),
      xaxis = list(title = "Year"),
      yaxis = list(title = "Average Score"),
      autosize = F, width = 1000, height = 500
    )
```

### Life Expectancy
```{r}
hap_df |>
  mutate(region = fct_reorder(region, health_life_expectancy))|>
  group_by(region, year)|>
  summarise(
    average_life_expectancy = mean(health_life_expectancy, na.rm = TRUE)
  ) |>
  plot_ly (
  x = ~region, y = ~average_life_expectancy, color = ~year,
  type = "bar", colors = "viridis") |>
  layout(
    title = "Healthy Life Expectancy By Region",
    xaxis = list(title = "Region"),
    yaxis = list(title = "Average Score"),
    autosize = F, width = 1000, height = 500
  )
```

**Regional Trend**
```{r}
regional_hap_df |>
  filter(factor == "healthy_life_expectancy")|>
  plot_ly(
      x = ~year, y = ~average_score, color = ~region,
      mode = 'lines+markers', type = 'scatter', colors = 'viridis',
      name = ~region) |>
    layout(
      title = paste("Regional Trend for Life Expectancy from 2015 to 2019"),
      xaxis = list(title = "Year"),
      yaxis = list(title = "Average Score"),
      autosize = F, width = 1000, height = 500
    )
```


### Freedom to Make Choice
```{r}
hap_df |>
  mutate(region = fct_reorder(region, freedom_to_make_life_choices))|>
  group_by(region, year)|>
  summarise(
    average_freedom = mean(freedom_to_make_life_choices, na.rm = TRUE)
  ) |>
  plot_ly (
  x = ~region, y = ~average_freedom, color = ~year,
  type = "bar", colors = "viridis") |>
  layout(
    title = "Freedom to Make Choice By Region",
    xaxis = list(title = "Region"),
    yaxis = list(title = "Average Score"),
    autosize = F, width = 1000, height = 500
  )
```

**Regional Trend**
```{r}
regional_hap_df |>
  filter(factor == "freedom_to_make_life_choices")|>
  plot_ly(
      x = ~year, y = ~average_score, color = ~region,
      mode = 'lines+markers', type = 'scatter', colors = 'viridis',
      name = ~region) |>
    layout(
      title = paste("Regional Trend for Freedom to Make Life Choices from 2015 to 2019"),
      xaxis = list(title = "Year"),
      yaxis = list(title = "Average Score"),
      autosize = F, width = 1000, height = 500
    )
```

### Perception of Corruption
```{r}
hap_df |>
  mutate(region = fct_reorder(region, perception_of_corruption))|>
  group_by(region, year)|>
  summarise(
    average_corruption = mean(perception_of_corruption, na.rm = TRUE)
  ) |>
  plot_ly (
  x = ~region, y = ~average_corruption, color = ~year,
  type = "bar", colors = "viridis") |>
  layout(
    title = "Perception of Corruption By Region",
    xaxis = list(title = "Region"),
    yaxis = list(title = "Average Score"),
    autosize = F, width = 1000, height = 500
  )
```

**Regional Trend**
```{r}
regional_hap_df |>
  filter(factor == "perception_of_corruption")|>
  plot_ly(
      x = ~year, y = ~average_score, color = ~region,
      mode = 'lines+markers', type = 'scatter', colors = 'viridis',
      name = ~region) |>
    layout(
      title = paste("Regional Trend for Perception of Corruption from 2015 to 2019"),
      xaxis = list(title = "Year"),
      yaxis = list(title = "Average Score"),
      autosize = F, width = 1000, height = 500
    )
```

### Generosity
```{r}
hap_df |>
  mutate(region = fct_reorder(region, generosity))|>
  group_by(region, year)|>
  summarise(
    average_generosity = mean(generosity, na.rm = TRUE)
  ) |>
  plot_ly (
  x = ~region, y = ~average_generosity, color = ~year,
  type = "bar", colors = "viridis") |>
  layout(
    title = "Generosity By Region",
    xaxis = list(title = "Region"),
    yaxis = list(title = "Average Score"),
    autosize = F, width = 1000, height = 500
  )
```

**Regional Trend**
```{r}
regional_hap_df |>
  filter(factor == "generosity")|>
  plot_ly(
      x = ~year, y = ~average_score, color = ~region,
      mode = 'lines+markers', type = 'scatter', colors = 'viridis',
      name = ~region) |>
    layout(
      title = paste("Regional Trend for Generosity from 2015 to 2019"),
      xaxis = list(title = "Year"),
      yaxis = list(title = "Average Score"),
      autosize = F, width = 1000, height = 500
    )
```

## World Map
```{r}
gdp <- worldjoin %>% 
  ggplot(mapping = aes(
    x=long,
    y=lat,
    group=group)) +
  scale_fill_viridis_c(option = "C") +
  coord_fixed(1.3) +
  geom_polygon(aes(fill=gdp_per_capita)) +
  ggtitle("GDP Per Capita") +
  cleanup

gdp
```




[Back To Home](index.html)

