---
title: "Regression Analysis"
author: "Tingyi Li"
date: "2023-12-03"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
set.seed(1)

knitr::opts_chunk$set(
  echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

#Part 1


#Multiple Linear Regression for Happiness Score Prediction

```{r}
happiness_score = read_csv("./Data/combined_happiness.csv")

predicted_model = lm(formula = score ~ gdp_per_capita + social_support + health_life_expectancy + freedom_to_make_life_choices + perception_of_corruption + generosity, 
                     data = happiness_score)

predicted_model|>
  broom::tidy()|>
  knitr::kable()

```

#Part 2

#Multiple linear Regression for Happiness Score Prediction in 2015
```{r}
data_2015 = happiness_score|>
  filter(year == 2015)

model2015 = lm(formula = score ~ gdp_per_capita + social_support + health_life_expectancy + freedom_to_make_life_choices + perception_of_corruption + generosity, 
                     data = data_2015)

model2015|>
  broom::tidy()|>
  knitr::kable()
```

#Multiple linear Regression for Happiness Score Prediction in 2016
```{r}
data_2016 = happiness_score|>
  filter(year == 2016)

model2016 = lm(formula = score ~ gdp_per_capita + social_support + health_life_expectancy + freedom_to_make_life_choices + perception_of_corruption + generosity, 
                     data = data_2016)

model2016|>
  broom::tidy()|>
  knitr::kable()
```

#Multiple linear Regression for Happiness Score Prediction in 2017
```{r}
data_2017 = happiness_score|>
  filter(year == 2017)

model2017 = lm(formula = score ~ gdp_per_capita + social_support + health_life_expectancy + freedom_to_make_life_choices + perception_of_corruption + generosity, 
                     data = data_2017)

model2017|>
  broom::tidy()|>
  knitr::kable()
```

#Multiple linear Regression for Happiness Score Prediction in 2018
```{r}
data_2018 = happiness_score|>
  filter(year == 2018)

model2018 = lm(formula = score ~ gdp_per_capita + social_support + health_life_expectancy + freedom_to_make_life_choices + perception_of_corruption + generosity, 
                     data = data_2018)

model2018|>
  broom::tidy()|>
  knitr::kable()
```

#Multiple linear Regression for Happiness Score Prediction in 2019
```{r}
data_2019 = happiness_score|>
  filter(year == 2019)

model2019 = lm(formula = score ~ gdp_per_capita + social_support + health_life_expectancy + freedom_to_make_life_choices + perception_of_corruption + generosity, 
                     data = data_2018)

model2019|>
  broom::tidy()|>
  knitr::kable()
```


#Part 3


