---
title: "Regression Analysis"
author: "Tingyi Li"
date: "2023-12-03"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
set.seed(1)

knitr::opts_chunk$set(
  echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

#Part 1


#Multiple Linear Regression for Happiness Score Prediction

```{r}
happiness_score = read_csv("./Data/combined_happiness.csv")

predicted_model = lm(formula = score ~ gdp_per_capita + social_support + health_life_expectancy + freedom_to_make_life_choices + perception_of_corruption + generosity, 
                     data = happiness_score)

predicted_model|>
  broom::tidy()|>
  knitr::kable()

```

The variable "freedom to make life choices" exhibits the largest coefficient (1.4812709) with a highly significant p-value, indicating it exerts the most substantial effect on happiness scores. Following this, "GDP per capita", "Health life expectancy", and "Social support" are also significant contributors to happiness in order of their estimates. "Perception of corruption" and "Generosity" also significantly impact happiness but to a lesser extent, based on their estimates and p-values.

#Part 2

#Multiple linear Regression for Happiness Score Prediction in 2015
```{r}
data_2015 = happiness_score|>
  filter(year == 2015)

model2015 = lm(formula = score ~ gdp_per_capita + social_support + health_life_expectancy + freedom_to_make_life_choices + perception_of_corruption + generosity, 
                     data = data_2015)

model2015|>
  broom::tidy()|>
  knitr::kable()
```

#Multiple linear Regression for Happiness Score Prediction in 2016
```{r}
data_2016 = happiness_score|>
  filter(year == 2016)

model2016 = lm(formula = score ~ gdp_per_capita + social_support + health_life_expectancy + freedom_to_make_life_choices + perception_of_corruption + generosity, 
                     data = data_2016)

model2016|>
  broom::tidy()|>
  knitr::kable()
```

#Multiple linear Regression for Happiness Score Prediction in 2017
```{r}
data_2017 = happiness_score|>
  filter(year == 2017)

model2017 = lm(formula = score ~ gdp_per_capita + social_support + health_life_expectancy + freedom_to_make_life_choices + perception_of_corruption + generosity, 
                     data = data_2017)

model2017|>
  broom::tidy()|>
  knitr::kable()
```

#Multiple linear Regression for Happiness Score Prediction in 2018
```{r}
data_2018 = happiness_score|>
  filter(year == 2018)

model2018 = lm(formula = score ~ gdp_per_capita + social_support + health_life_expectancy + freedom_to_make_life_choices + perception_of_corruption + generosity, 
                     data = data_2018)

model2018|>
  broom::tidy()|>
  knitr::kable()
```

#Multiple linear Regression for Happiness Score Prediction in 2019
```{r}
data_2019 = happiness_score|>
  filter(year == 2019)

model2019 = lm(formula = score ~ gdp_per_capita + social_support + health_life_expectancy + freedom_to_make_life_choices + perception_of_corruption + generosity, 
                     data = data_2018)

model2019|>
  broom::tidy()|>
  knitr::kable()
```


#Part 3

#Multiple linear Regression for Happiness Score Prediction in Central and Eastern Europe
```{r}
central_east = happiness_score|>
  filter(region == "Central and Eastern Europe")

model1 = lm(formula = score ~ gdp_per_capita + social_support + health_life_expectancy + freedom_to_make_life_choices + perception_of_corruption + generosity, 
                     data = central_east)

model1|>
  broom::tidy()|>
  knitr::kable()
```


#Multiple linear Regression for Happiness Score Prediction in Eastern Asia
```{r}
east_asia = happiness_score|>
  filter(region == "Eastern Asia")

model2 = lm(formula = score ~ gdp_per_capita + social_support + health_life_expectancy + freedom_to_make_life_choices + perception_of_corruption + generosity, 
                     data = east_asia)

model2|>
  broom::tidy()|>
  knitr::kable()
```

#Multiple linear Regression for Happiness Score Prediction in Latin America and Caribbean
```{r}
latin_caribbean = happiness_score|>
  filter(region == "Latin America and Caribbean")

model3 = lm(formula = score ~ gdp_per_capita + social_support + health_life_expectancy + freedom_to_make_life_choices + perception_of_corruption + generosity, 
                     data = latin_caribbean)

model3|>
  broom::tidy()|>
  knitr::kable()
```


#Multiple linear Regression for Happiness Score Prediction in Middle East and Northern Africa
```{r}
east_africa = happiness_score|>
  filter(region == "Middle East and Northern Africa")

model4 = lm(formula = score ~ gdp_per_capita + social_support + health_life_expectancy + freedom_to_make_life_choices + perception_of_corruption + generosity, 
                     data = east_africa)

model4|>
  broom::tidy()|>
  knitr::kable()
```


#Multiple linear Regression for Happiness Score Prediction in North America
```{r}
north_america = happiness_score|>
  filter(region == "North America")

model5 = lm(formula = score ~ gdp_per_capita + social_support + health_life_expectancy + freedom_to_make_life_choices + perception_of_corruption + generosity, 
                     data = north_america)

model5|>
  broom::tidy()|>
  knitr::kable()
```

#Multiple linear Regression for Happiness Score Prediction in Southeastern Asia
```{r}
southeastern_asia = happiness_score|>
  filter(region == "Southeastern Asia")

model6 = lm(formula = score ~ gdp_per_capita + social_support + health_life_expectancy + freedom_to_make_life_choices + perception_of_corruption + generosity, 
                     data = southeastern_asia)

model6|>
  broom::tidy()|>
  knitr::kable()
```

#Multiple linear Regression for Happiness Score Prediction in Southern Asia
```{r}
southern_asia = happiness_score|>
  filter(region == "Southern Asia")

model7 = lm(formula = score ~ gdp_per_capita + social_support + health_life_expectancy + freedom_to_make_life_choices + perception_of_corruption + generosity, 
                     data = southern_asia)

model7|>
  broom::tidy()|>
  knitr::kable()
```

#Multiple linear Regression for Happiness Score Prediction in Sub-Saharan Africa
```{r}
sub_africa = happiness_score|>
  filter(region == "Sub-Saharan Africa")

model8 = lm(formula = score ~ gdp_per_capita + social_support + health_life_expectancy + freedom_to_make_life_choices + perception_of_corruption + generosity, 
                     data = sub_africa)

model8|>
  broom::tidy()|>
  knitr::kable()
```

#Multiple linear Regression for Happiness Score Prediction in Western Europe
```{r}
west_europe = happiness_score|>
  filter(region == "Western Europe")

model9 = lm(formula = score ~ gdp_per_capita + social_support + health_life_expectancy + freedom_to_make_life_choices + perception_of_corruption + generosity, 
                     data = west_europe)

model9|>
  broom::tidy()|>
  knitr::kable()
```

